IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-10-18 paper/appendix.tex

In the main body of the paper we have only sketched the proof structure
for Lemma~\ref{lemma:rr-tsni-general}.  Here we give it with full
details, and first state and proof another lemma that will be used in that
proof.

@@

%\begin{lemma}
%  \label{lemma:erase-context}
%  For any expressions |ie1|, |ie2|, |ie1'|, context |iE1| as well as
%  label |il|, if
%  \[ |ie1= iHole2 iE1 ie1'| \quad \land \quad |erase il ie1 = erase il ie2| \]
%  then there exist |iE2|, |ie2'| such that
%  \[ |ie2= iHole2 iE2 ie2'| \quad \land \quad
%  |erase il ie1' = erase il ie2'|\neq|bullet| \]
%  Similarly, for any
%  expressions |ie1|, |ie2|, |te1'|, context |iE1| as well as
%  label |il|, if
%  \[ |ie1= tHole2 iE1 te1'| \quad \land \quad |erase il ie1 = erase il ie2| \]
%  then there exist |iE2|, |te2'| such that
%  \[ |ie2= tHole2 iE2 te2'| \quad \land \quad
%  |erase il te1' = erase il te2'|\neq|bullet| \]
%\end{lemma}
%\begin{proof}
%  TODO
%\end{proof}

\begin{proof}[Proof of Lemma~\ref{lemma:rr-tsni-general}]
  First, we observe there must be at least one task in |ic1|, otherwise
  it could not take a step.  Thus, |ic1| is of the form
  |iconf iS1 (it1, its1)|.
  Furthermore, let |ic2| be |iconf iS2 its2|.
  Consider two cases:
  \begin{itemize}
    \item $|erase il it1|=|bullet|$.
    By the definition of |erasef il|, we know that |il canFlowTo lcurr|
    where |lcurr| is the label of |it1|.
    In this case, we do not need to take a step for
    |ic2|, because |ic2'=ic2| will already be |il|-equivalent to |ic1'|.
    To show this, note that the tasks |its1| in |ic1| are left in the
    same order and unmodified (the scheduling policy only
    modifies the first task). The task |it1| either
    gets dropped (by \textsc{I-noStep}), or
    transforms into a task |it1'| as well as potentially spawning a new
    task |it1''|.  Since both |it1'| and |it1''| have a label that is
    at least as high as the label of |it1| (can be seen
    by inspecting all reduction rules), they will get filtered
    by |erasef il| in |ic1'|.  Therefore, the |il|-equivalence of the
    task list is guaranteed.
    Lets consider the possible changes to |iS1|:
    Only five reduction interact with |iS1|,
    thus it suffices to consider these cases:
    \begin{description}
      \item[Case \textsc{I-send}]
      A new message triple with label |il'| gets added to the message
      queue of the receiving thread.  However, since |lcurr canFlowTo il'|,
      the triple will get erased.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      In this case, only the queue of
      task |it1| can change, which gets erased.
      \item[Case \textsc{I-new}] The newly allocated address has to be at a
      label at least as high as |lcurr|, and will thus be erased.
      \item[Case \textsc{I-write}] Only addresses with a label |il'| above
      |lcurr| can be written, thus the change in |iS1| will get erased.
    \end{description}
    This ensures that $|ic1'|\approx_{|il|}|ic2'|$, as well as
    $|ic2| |.->|^* |ic2'|$ (in zero steps), as claimed.
    %    \alphacondition{We need all scheduling policies to not change the order
    %      of any tasks (except for the first one).  Newly spawned task can appear
    %      anywhere in the list.}
    \item $|erase il it1|\neq|bullet|$.
    By the definition of |erasef il|, the task list |its2|
    in |ic2| must be of the
    form |its2', it2, its2''| (for some task lists |its2'|, |its2''| and
    some task |it2|)
    where
    \begin{align}
    |erase il its2'| = |nil| \\
    |erase il it2| = |erase il it1| \\
    |erase il its2''| = |erase il its1|
    \end{align}
    (where |nil| is the empty list of tasks).
    Now, intuitively we will first execute a number of steps to process
    the tasks in |its2'| (execute them one step and move them to the back
    of the task list, or drop them if they are done or stuck).  Then, the task
    |it2| can take the same step as |it1|, which will result in a configuration
    |ic2'|
    with the desired properties.
    More formally, we can proceed as follows:

    Now, the first task |it2| in |ic2''| is |il|-equivalent to the task |it1|.
    This implies that the two tasks must have the same id, label and
    can only differ in the expression or store if some subexpression
    is of the form |Labeled il' ie|.  In this case, the expression |ie| could
    be different in the two threads if |lcurr canFlowTo il'|.  However, none of the reduction rules
    depend on an expression in that position, and there is never a
    hole in that position
    where evaluation could take place.  Thus, the same rules will syntactically
    match for both task, and we are left to argue that all premises
    evaluate to the same values for |it1| and |it2|, as well as that
    the resulting states |iS1'| and
    |iS2''| are |il|-equivalent.
    The additional premises $P$ that follow
    the condition in Definition~\ref{def:restricted} are not a problem,
    since those
    predicates only depend on |erase il ic1|, which is equivalent
    to |erase il ic2''|, and thus those predicates evaluate in the same way.
    All other premises are either on the threads labels (which are the same),
    or on the state |iS1|, or |iS2'|, respectively.  Because
    |erase il iS1=erase il iS2'|, all of these also evaluate in the same way,
    as can be seen by simply considering all rules that involve or
    change the state:
    \begin{description}
      \item[Case \textsc{I-send}]
      Here, the task |it2| will send the same message to the same
      receiver queue. This
      queue is either completely erased, or it is |il|-equivalent.  In both
      cases, |il|-equivalence of |iS1'| and |iS2'| is preserved.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      When the tasks are receiving a message, then by the reduction rules
      we know that they first filter the queue by the label
      |lcurr| of |it1|.  We
      also know that the queues are equivalent when filtered by the less
      restrictive label |il|, thus the messages received (or dropped) from the
      queue are equivalent.
      \item[Case \textsc{I-new}] The newly allocated address can be the same
      for both |it1| and |it2|, thus resulting in |il|-equivalent states.
      \item[Case \textsc{I-write}] By |erase il it1=earse il it2| both tasks
      write the same value, and therefore the resulting states will still
      be |il|-equivalent.
    \end{description}
    After |it2| has taken a step, we finally arrive in the desired
    configuration |ic2'=iconf iS2'' (its2'', its2''', its2'''')|, where
    |its2''''| contains the task resulting from executing |it2| (and
    might contain, zero (if the task was done or stuck), one (for most steps) or two tasks if a new task was launched).
    As required, we have
    \[ |ic2| |.->|^* |ic2''| |.->| |ic2'|
    \quad \land \quad |ic1'| \approx_{|il|} |ic2'| \text{.}\]
  \end{itemize}
\end{proof}

>>>>

In this section we prove the theorems we have stated in the paper.
We first observe that the non-interference claims for the languages
|specLang seqf targetLang| and |specLang roundrobinf targetLang|
in Theorems~\ref{thm:rr-tsni} and~\ref{thm:seq-tini} follow directly
from Theorem~\ref{thm:restricted}, where the set
of predicates is the set of always valid predicates (i.e., no restriction).

Before we proceed with the proof of Theorem~\ref{thm:restricted},
we state and proof two lemmas we will use.

\begin{lemma}
  \label{lemma:rr-tsni-general}
  We consider, for any target language |targetLang|,
  the restricted IFC language |restrictedLang alpha targetLang|
  (according to Definition~\ref{def:restricted}).
  Then,
  for any configurations |ic1|, |ic1'|, |ic2|, and label |il| where
  \begin{equation} \label{eq:tsni-lemma-lhs}
  |ic1| \approx_{|il|} |ic2|
  \qquad \text{and} \qquad
  |ic1| |.->| |ic1'|
  \end{equation}
  there exists a configuration |ic2'| such that
  \begin{equation} \label{eq:tsni-lemma-rhs}
  |ic1'| \approx_{|il|} |ic2'|
  \qquad \text{and} \qquad
  |ic2| |.->|^* |ic2'|
  \ \text{.}
  \end{equation}
\end{lemma}
\begin{proof}
  First, we observe there must be at least one task in |ic1|, otherwise
  it could not take a step.  Thus, |ic1| is of the form
  |iconf iS1 (it1, its1)|.
  Furthermore, let |ic2| be |iconf iS2 its2|.
  Consider two cases:
  \begin{itemize}
    \item $|erase il it1|=|bullet|$.
    By the definition of |erasef il|, we know that |il canFlowTo lcurr|
    where |lcurr| is the label of |it1|.
    In this case, we do not need to take a step for
    |ic2|, because |ic2'=ic2| will already be |il|-equivalent to |ic1'|.
    To show this, note that the tasks |its1| in |ic1| are left in the
    same order and unmodified (the scheduling policy only
    modifies the first task). The task |it1| either
    gets dropped (by \textsc{I-noStep}), or
    transforms into a task |it1'| as well as potentially spawning a new
    task |it1''|.  Since both |it1'| and |it1''| have a label that is
    at least as high as the label of |it1| (can be seen
    by inspecting all reduction rules), they will get filtered
    by |erasef il| in |ic1'|.  Therefore, the |il|-equivalence of the
    task list is guaranteed.
    Lets consider the possible changes to |iS1|:
    Only five reduction interact with |iS1|,
    thus it suffices to consider these cases:
    \begin{description}
      \item[Case \textsc{I-send}]
      A new message triple with label |il'| gets added to the message
      queue of the receiving thread.  However, since |lcurr canFlowTo il'|,
      the triple will get erased.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      In this case, only the queue of
      task |it1| can change, which gets erased.
      \item[Case \textsc{I-new}] The newly allocated address has to be at a
      label at least as high as |lcurr|, and will thus be erased.
      \item[Case \textsc{I-write}] Only addresses with a label |il'| above
      |lcurr| can be written, thus the change in |iS1| will get erased.
    \end{description}
    This ensures that $|ic1'|\approx_{|il|}|ic2'|$, as well as
    $|ic2| |.->|^* |ic2'|$ (in zero steps), as claimed.
    %    \alphacondition{We need all scheduling policies to not change the order
    %      of any tasks (except for the first one).  Newly spawned task can appear
    %      anywhere in the list.}
    \item $|erase il it1|\neq|bullet|$.
    By the definition of |erasef il|, the task list |its2|
    in |ic2| must be of the
    form |its2', it2, its2''| (for some task lists |its2'|, |its2''| and
    some task |it2|)
    where
    \begin{align}
    |erase il its2'| = |nil| \\
    |erase il it2| = |erase il it1| \\
    |erase il its2''| = |erase il its1|
    \end{align}
    (where |nil| is the empty list of tasks).
    Now, intuitively we will first execute a number of steps to process
    the tasks in |its2'| (execute them one step and move them to the back
    of the task list, or drop them if they are done or stuck).  Then, the task
    |it2| can take the same step as |it1|, which will result in a configuration
    |ic2'|
    with the desired properties.
    More formally, we can proceed as follows:

    Now, the first task |it2| in |ic2''| is |il|-equivalent to the task |it1|.
    This implies that the two tasks must have the same id, label and
    can only differ in the expression or store if some subexpression
    is of the form |Labeled il' ie|.  In this case, the expression |ie| could
    be different in the two threads if |lcurr canFlowTo il'|.  However, none of the reduction rules
    depend on an expression in that position, and there is never a
    hole in that position
    where evaluation could take place.  Thus, the same rules will syntactically
    match for both task, and we are left to argue that all premises
    evaluate to the same values for |it1| and |it2|, as well as that
    the resulting states |iS1'| and
    |iS2''| are |il|-equivalent.
    The additional premises $P$ that follow
    the condition in Definition~\ref{def:restricted} are not a problem,
    since those
    predicates only depend on |erase il ic1|, which is equivalent
    to |erase il ic2''|, and thus those predicates evaluate in the same way.
    All other premises are either on the threads labels (which are the same),
    or on the state |iS1|, or |iS2'|, respectively.  Because
    |erase il iS1=erase il iS2'|, all of these also evaluate in the same way,
    as can be seen by simply considering all rules that involve or
    change the state:
    \begin{description}
      \item[Case \textsc{I-send}]
      Here, the task |it2| will send the same message to the same
      receiver queue. This
      queue is either completely erased, or it is |il|-equivalent.  In both
      cases, |il|-equivalence of |iS1'| and |iS2'| is preserved.
      \item[Case \textsc{I-recv} and \textsc{I-noRecv}]
      When the tasks are receiving a message, then by the reduction rules
      we know that they first filter the queue by the label
      |lcurr| of |it1|.  We
      also know that the queues are equivalent when filtered by the less
      restrictive label |il|, thus the messages received (or dropped) from the
      queue are equivalent.
      \item[Case \textsc{I-new}] The newly allocated address can be the same
      for both |it1| and |it2|, thus resulting in |il|-equivalent states.
      \item[Case \textsc{I-write}] By |erase il it1=earse il it2| both tasks
      write the same value, and therefore the resulting states will still
      be |il|-equivalent.
    \end{description}
    After |it2| has taken a step, we finally arrive in the desired
    configuration |ic2'=iconf iS2'' (its2'', its2''', its2'''')|, where
    |its2''''| contains the task resulting from executing |it2| (and
    might contain, zero (if the task was done or stuck), one (for most steps) or two tasks if a new task was launched).
    As required, we have
    \[ |ic2| |.->|^* |ic2''| |.->| |ic2'|
    \quad \land \quad |ic1'| \approx_{|il|} |ic2'| \text{.}\]
  \end{itemize}
\end{proof}
With this, it is easy to proof Theorem~\ref{thm:restricted} as follows.
\begin{proof}[Proof of Theorem~\ref{thm:restricted}, TSNI]
% short version:
%  We prove the theorem by induction on the length of the derivation sequence in~\eqref{eq:tsni-lhs}.
%  The base case for derivations
%  of length 0 is trivial, allowing
%  us to simply chose $|ic2'=ic2|$.  In the step case,
%  we split the derivation sequence from~\eqref{eq:tsni-lhs} of length $n+1$ as
%  $|ic1| |.->| |ic1''| |.->|^n |ic1'|$
%  for some configuration |ic1''|.  By Lemma~\ref{lemma:rr-tsni-general} to
%  the first step and
%  the induction hypothesis to the rest of the derivation sequence,
%  we directly get the desired property.
  We proof the theorem by induction on the length of the derivation sequence in~\eqref{eq:tsni-lhs}.
  The base case for derivations
  of length 0 is trivial, allowing
  us to simple chose $|ic2'=ic2|$.  In the step case, we assume
  the theorem holds for derivation sequences of length up to $n$, and show that it also
  holds for those of length $n+1$.  We split the derivation sequence from~\eqref{eq:tsni-lhs} as follows:
  \[
  |ic1| |.->| |ic1''| |.->|^n |ic1'|
  \]
  for some configuration |ic1''|.  By Lemma~\ref{lemma:rr-tsni-general}, we get
  |ic''| with
  \begin{equation} \label{eq:tsni-proof-1}
  |ic1''| \approx_{|il|} |ic2''|
  \qquad \text{and} \qquad
  |ic2| |.->|^* |ic2''|
  \end{equation}
  Applying the induction hypothesis to
  $|ic1''| |.->|^n |ic1'|$, we get |ic2'| with
  \begin{equation} \label{eq:tsni-proof-2}
  |ic1'| \approx_{|il|} |ic2'|
  \qquad \text{and} \qquad
  |ic2''| |.->|^* |ic2'|
  \end{equation}
  Stitching together the derivation sequences from~\eqref{eq:tsni-proof-1} and~\eqref{eq:tsni-proof-2} directly gives
  us the right-hand side of the implication in the TSNI
  definition~\eqref{eq:tsni-rhs}, which concludes the proof.
\end{proof}

@@
